%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title (en): Multiagent Systems and Organizations                           %%
%% Title (cs): Multiagentní systémy a organizace                              %%
%%                                                                            %%
%% Author: Bc. Lukáš Kúdela                                                   %%
%% Supervisor: Prof. RNDr. Petr Štěpánek, DrSc.                               %%
%%                                                                            %%
%% Academic year: 2011/2012                                                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Supervisor
First I would like to thank my supervisor, prof. RNDr. Petr Štěpánek for sparking my interest in multiagent systems, coming up with a interesting and challenging topic and for the time we have spent together discussing matters more or less related to the subject of this thesis.
% Parents
I would also like to say that I am thankful to my parents, Libuša and Ivan, who have always believed in the importance of quality education and made their best to ensure their children receive the best one possible.
I am also very grateful to them for never giving me a reason to worry about issues that would prevent me from fully concentrating on my studies, which they supported in every possible way.
% Jakub
Next, I would like express my gratitude to my brother, Jakub, who has helped me edit some of the images in this thesis.
% Monika
Finally, I would like to thank Monika, who has been bearing all the consequences of having a thesis-writing boyfriend and has managed not to lose her nerve.
Once again, thank you all!

% Temporarily select the Slovak language (with French spacing).
\selectlanguage{slovak}
\frenchspacing

% Vedúci
Najpr by som chcel poďakovať môjmu vedúcemu, prof. RNDr. Petrovi Štěpánkovi, za vzbudenie môjho záujmu o multiagentové systémy, navrnutie zaujímavej témy, ktorá bola výzvou a za čas, ktorý sme spoločne strávili diskutovaním o veciach viac či menej súvisiacích s predmetom tejto práce.
% Rodičia
Chcel by som taktiež povedať, že som vďačný svojím rodičom, Ivanovi a Libuši, ktorí vždy verili, že kvalitné vzdelanie je dôležité a urobili všetko preto, aby ich deti dostali to najlepšie možné.
Som im taktiež vďačný za to, že mi nikdy nedali dôvod trápiť sa záležitosťami, ktoré by mi zabránili plne sa sústrediť na štúdium, ktoré podporovali vo všetkých ohľadoch.
% Jakub
Ďalej by som chcel vyjadriť vďačnosť môjmu bratovi, Jakubovi, ktorý mi pomohol upraviť niektoré z obrázkov v tejto práci.
% Monika
Napokon by som chcel poďakovať Monike, ktorá znášala všetky následky toho, keď jej priateľ písal túto prácu a podarilo sa jej pri tom nestratiť nervy.
Ešte raz, ďakujem všetkým!

% Select the English language (with English spacing) again.
\selectlanguage{english}
\nonfrenchspacing 