%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title (en): Multiagent Systems and Organizations                           %%
%% Title (cs): Multiagentní systémy a organizace                              %%
%%                                                                            %%
%% Author: Bc. Lukáš Kúdela                                                   %%
%% Supervisor: Prof. RNDr. Petr Štěpánek, DrSc.                               %%
%%                                                                            %%
%% Academic year: 2011/2012                                                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Title: Multiagent Systems and Organizations\\
Author: Bc. Lukáš Kúdela\\
Author's e-mail address: \url{lukas.kudela@gmail.com}\\
Department: Department of Theoretical Computer Science and Mathematical Logic\\
Thesis Supervisor: Prof. RNDr. Petr Štěpánek, DrSc.\\
Supervisor's e-mail address: \url{petr.stepanek@mff.cuni.cz}\\

% Motivation - Why do we care about the problem and the results?
Abstract: Multiagent systems (MASs) are emerging as a promising paradigm for conceptualizing, designing and implementing large-scale heterogeneous software systems.
The key advantage of looking at components in such systems as autonomous agents is that as agents they are capable of flexible self-organization, instead of being rigidly organized by the system's architect.
However, self-organization is like evolution -- it takes a lot of time and the results are not guaranteed.
More often than not, the system's architect has an idea about how the agents should organize themselves -- what types of organizations they should form.
% Problem statement - What problem are we trying to solve?
In our work, we tried to solve the problem of modelling organizations and their roles in a MAS, independent on the agent platform on which the MAS will eventually run.
% Approach - How did we go about solving or making progress on the problem?
First and foremost, we have proposed a metamodel for expressing platform-independent organization models.
Furthermore, we have implemented the proposed metamodel for the Jade agent platform as a module extending the framework.
Finally, we have demonstrated the usage of this module by modelling three specific organization types: remote function invocation, arithmetic expression evaluation and sealed-bid auction.
% Results - What's the answer?
Our work shows how to separate the acquired role-related behaviour from the behaviour intrinsic to an agent. 
% Conclusions - What are the implications of our answer?   
This separation enables organizations to be developed independently of the agents that will participate in them, thus facilitating the development of open systems.

Keywords: multiagent systems, organizations, roles, metamodel