%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% MASTER'S THESIS                                                            %%
%%                                                                            %% 
%% Title (en): Multi-Agent Systems and Organizations                          %%
%% Title (cs): Multiagentní systémy a organizace                              %%
%%                                                                            %%
%% Author: Bc. Lukáš Kúdela                                                   %%
%% Supervisor: Prof. RNDr. Petr Štěpánek, DrSc.                               %%
%%                                                                            %%
%% Academic year: 2011/2012                                                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Platform-Independent Metamodel---Thespian}

% Chapter intro
In this chapter, we will present our platform-independent metamodel for modelling organizations in MASs---\textit{Thespian}.

% Thespian - defnition
\textit{Thespian} is a metamodel to which platform-independent models of organizations in MASs must conform; in more detail, it is a model\footnote{Recall that a metamodel is a kind of model.} representing a modelling language \footnote{When can call this language the \textit{Thespian modelling language}.} to which platform-independent models of organizations in MASs must conform.

% Thespian name inspiration - Thespis of Icaria
\textit{Thespian} is named after \textit{Thespis of Icaria} (present-day Dionysos, Greece), who lived in the 6th century BC and, according to certain Ancient Greece sources and especially Aristotle, was the first person ever to appear on stage as an actor playing a character in a play (instead of just speaking as himself) \cite{Wikipedia-Thespis}.

% Thespian inspiration - Aalaadin, O&P, PIM4Agents and powerJade
\textit{Thespian} is inspired by all four metamodels introduced in the previous chapter: \textit{Aalaadin}, \textit{O\&P}, \textit{PIM4Agents} and \textit{powerJade}.
% Aaalaadin
Similarly to \textit{Aalaadin}, it contains both dynamic and static concepts (called \textit{core} and \textit{methodological} concepts in Aalaadin).
Instances of dynamic/static concepts will end up as run-time/design-time entities in the platform-specific model.
% O&P
Like \textit{O\&P}, it can be used to model holonic MASs.
% PIM4Agents
Similarly to \textit{PIM4Agents}, it enables explicit modelling of protocols and messages exchanged in these protocols.
% powerJade 
And finally like \textit{powerJade}, it is able to represent competences and responsibilities (called \textit{powers} and \textit{requirements} respectively in \textit{powerJade}) of roles.

% Two partitions: {organization, player and protocol}, {desgn-time and run-time}
The \textit{Thespian} metamodel can be partitioned in two orthogonal ways: in a high-level and low-level way.
Both partitions will be described in the following sections; the integrated metamodel will be presented in the last section.

% High-level partition
The high-level partition divides the concepts according to the area they represent: organization, player or protocol.
This is a more natural partition of the two since fewer dependencies among concepts from different parts exists than in the other partition.
The organization/protocol part of the MAS can be designed (and implemented) independently of the player part; indeed agent developed by one team can become members of organizations and play roles developed by another team. 

% Low-level partition
The low-level partition separates the concepts based on whether they represent design-time or run-time entities.
The design-time entities are created already at design-time and are usually implemented as \textit{agent classes} in the target agent platform; they are analogous to the concept of \textit{Class} in OOP.
The run-time entities are created only at run-time and are usually implemented as \textit{agent instances} in the target agent platform; they are analogous to the concept of \textit{Instance} in OOP.

% Type-token distinction - definition
Before continuing with the presentation of \textit{Thespian}, it is important to explain the \textit{type-token distinction}.
In disciplines such as philosophy and knowledge representation, the type-token distinction is a distinction that separates a \textit{concept} from objects that are particular \textit{instances} of that concept.
\textit{Thespian} has been designed to support the type-token distinction and the correct differentiation is a recurring theme in this chapter.

% Type-token distinction - MAS
As an example of the type-token distinction, consider a MAS.
The \textit{specification} of the MAS (its source code) is the \textit{MAS type} and its \textit{manifestation} (a running MAS) is the \textit{MAS token}.
Just like a type can (and usually does) have many tokens, a MAS specification can have multiple manifestations---the same source code can be run many times.

% Static & dynamic metamodels
The \textit{Thespian} metamodel includes both a static and dynamic metamodels for describing structural and behavioural aspects of organizations respectively.
In the following two sections, we will introduce both metamodels in detail.

\input{41-ch4-section1-static-metamodel.tex}

\input{42-ch4-section2-dynamic-metamodel.tex}